The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-10-02 paper2_JNLS/paper.tex
NM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the region of shallower depth, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical foundation that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019anomalous} and Majda et al.~\cite{majda2019statistical, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, in particular on the relationship between surface-displacment skewness and surface-slope variance. Comparison between the two further confirms the predictive power of the theoretical framework developed in \cite{majda2019statistical}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly discussed in \cite{majda2019statistical} for the sake of brevity.

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and the Benjamin-Feir deep-water modulational instability. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events occur much more frequently when surface statistics deviate from Gaussian. In this way, it makes sense to approach anomalous waves from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine}.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics from a randomized wave-field encountering an abrupt depth change (ADC). In these experiments, the incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the shallower region, the wave distribution skews strongly towards positive displacement, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} developed a theoretical framework that accurately captures several key aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi \cite{majdaqi2019} analyzed more severely truncated systems of lower dimension that were shown to exhibit many of the same anomalous properties while enjoying a more tractable structure.

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019anomalous} and Majda et al.~\cite{majda2019statistical, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, particularly regarding a relationship between surface-displacement skewness and surface-slope variance that emerges from the theory. Comparison with the new measurements further confirms the predictive power of the theoretical framework developed by Majda et al.~\cite{majda2019statistical}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.
